---
title: "CAE Cubic Fit"
author: "Matt Brauer"
date: '`r format(Sys.Date(), "%B %d, %Y")`'
output:
  BiocStyle::html_document:
    highlight: pygments
    number_sections: no
    theme: united
    toc: no
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
source("functions.R")
source("data.R")
```
## Remember the "Cubic model"?

In early May the White House Council of Economic Advisors (CEA), led by the odious and always wrong Kevin Hasset put out a "model" of covid-19 deaths that resembled the output of a fudged excel curve-smoothing. This "cubic model" was cited as evidence that the pandemic would soon be over, and was contrasted with the projections from the ACTUAL model developed by IHME and run in May 4.

Here is what the author of "Dow 36,000" projected for Covid-19 deaths:

```{r cubic, echo=FALSE, warning=FALSE, message=FALSE}
library(knitr)
imgpath <- "IHME/cubic_model.jpeg"
img <- readJPEG(imgpath)
knitr::include_graphics(imgpath)
```
Note that the May 4 IHME model is the most extreme, and that it was portrayed by the White House as hopelessly pessimistic.

How has that May 4 model fared?

```{r ihme-vs-cubic, echo=FALSE, warning=FALSE, message=FALSE, fig.width = 12, fig.cap="Predicted vs. actual deaths. Prediction from May 5 2020 run of UW/IHME model."}
csse_by_country %>%
  filter(country == "US") %>%
  dplyr::rename(actual = deaths) %>%
  left_join(ihme_05_04) %>%
  daily_from_cumulative() %>%
  select(date, new_actual, new_ihme) %>%
  dplyr::rename(actual = new_actual,
                IHME = new_ihme) %>%
  filter(!is.na(actual), !is.na(IHME)) %>%
  ungroup() %>%
  pivot_longer(-date, names_to="source", values_to="deaths") %>%
  ggplot(aes(x=date, y=deaths, color=source)) +
#  xlim("2020-03-01", NA) +
  geom_line() +
  theme_minimal()

```

Sadly, it's been a bit optimistic, but mostly spot on.

Kevin Hasset has since left the CEA, looking for new rubes to confuzzle with his mad economist's Excel skillz.

## Data sources

* [JHU dashboard](https://coronavirus.jhu.edu/map.html)
* [UW Institute for Health Metrics and Evaluation (IHME)](http://www.healthdata.org/covid/data-downloads)

